---
title: "Assignment 1"
author: "Group 26: Aoming Sun,Siran Shen,Tong Pei"
output:
  pdf_document:
    extra_dependencies: ["float", "caption","booktabs"]
    fig_caption: true
    keep_tex: yes
geometry: "left=2cm,right=2cm,top=1.5cm,bottom=1.5cm"
fontsize: 6pt
highlight: tango
header-includes:
    \usepackage{float} 
    \floatplacement{figure}{H}  
    \floatplacement{table}{H} 
---



_In order not to be bothered with rounding the numbers, set 
`options(digits=3)` `r options(digits=3)`._


## Exercise 1
**a)**
```{r include=FALSE}
library(patchwork)
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
library(broom)
library(kableExtra)
```
To test the normality of the dataset we have plotted 2 qqplot for both before and after the replacement by margarine.
```{r Assignment_1a, echo=FALSE, fig.cap="Q-Q plots of cholesterol levels before and after the diet.", fig.pos="H", fig.height=2.5, fig.width=5, fig.align='center'}

data <- read.table(file = "cholesterol.txt",header=TRUE)
col_b <- data$Before
col_a <- data$After8weeks

plot_before <- ggplot(data, aes(sample = Before)) +
  stat_qq(color = "blue") +
  stat_qq_line(color = "red") +
  labs(title = "Before Diet", x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_minimal()


plot_after <- ggplot(data, aes(sample = After8weeks)) +
  stat_qq(color = "darkgreen") +
  stat_qq_line(color = "red") +
  labs(title = "After 8 Weeks", x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_minimal()


combined_plot <- plot_before + plot_after
print(combined_plot)

```


Figure1 showed that the data both before and after the experiment is distributed normally.

Since data are normally distributed we used pearson correlation coefficient.

```{r,fig.pos="H"}
result_cor <- cor.test(col_b, col_a, method = "pearson")
```

```{r, echo=FALSE, fig.env="figure", fig.pos="!ht"}
tidy_result_cor <- broom::tidy(result_cor)


knitr::kable(tidy_result_cor,
             caption = "Pearson correlation of before and after diet.",float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")

```
With this pretty low p value and the high correlation coefficient 
we can say the data of the columns are highly correlated. 

**b)**
First of all, we believe the dataset is a paired dataset because the data of the same row is 
from the same participant before and after the experiment. 
Therefore Mann Whitney test is not applicable since it's not for paired data test 
and permutation test is applicable.

T-test: Since the data is normal and we are expecting a decline after the experiment, 
we can use the mean value of the cholesterol level as the static. Therefore we choose the t test.
```{r}
result_tt <- t.test(col_b, col_a, paired = TRUE, alternative = "greater")
```

```{r, echo=FALSE, fig.env="figure", fig.pos="!htb"}
tidy_result_tt <- broom::tidy(result_tt)


knitr::kable(tidy_result_tt,
             caption = "T-test to to verify whether the diet with low fat margarine has an eﬀect.",float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")
```
The data were paired, so the paired t-test and Wilcoxon signed rank test were applied.
```{r}
# Wilcoxon
wilcox_test <- wilcox.test(data$Before, data$After8weeks, paired = TRUE)
```
```{r echo=FALSE}
tidy_wilcox_test <- broom::tidy(wilcox_test)


knitr::kable(tidy_wilcox_test,
             caption = "Wilcoxon test to to verify whether the diet with low fat margarine has an eﬀect.",float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")
```
The P value of both tests are very small so we confidently reject the H0 hypothesis with 95\% confidence 
to say the drop of cholesterol average level is at least 0.556.


Permutation test: Permutation test can directly give the significance of the change in the data.
```{r include=FALSE}
library(perm)
```

```{r}
permTS(col_b,col_a, alternative = "greater", method = "exact.mc",control = permControl(nmc = 10000))
```
```{r}
mystat = function(x, y) {
  mean(x - y)
}
B = 10000
tstar = numeric(B)
pl = 0
pr = 0
for (i in 1: B) {
  pairstar = t(apply(cbind(col_b, col_a), 1, sample))
  # print(pairstar)
  tstar[i] = mystat(pairstar[, 1], pairstar[, 2])
}
myt = mystat(col_b, col_a)
print(sum(tstar > myt))
pl = sum(tstar < myt) / B
pr = sum(tstar > myt) / B
p = 2 * min(pl, pr)
print(p)
```
After the random swap of the data before and after the experiment, 
the mean difference never reached 0.62 again thus p = 0 and we can reject the H0 confidently .

Mann-Whitney test is not applicable as it applies to independent samples.


**c)**
To find the mean value of the column after we use the t-test because it is a normal distribution.
```{r, fig.height=3, fig.width=6}
#c1
col_a_tt <- t.test(col_a, conf.level = 0.97)
```
```{r, echo=FALSE, fig.env="figure", fig.pos="!htb"}
tidy_col_a_tt <- broom::tidy(col_a_tt)


knitr::kable(tidy_col_a_tt,
             caption = "T-test of After8weeks data .",float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")
```
With 97\% confidence, we can say the mean value of the column after is between (5.164, 6.394)

Bootstrap test:
```{r, fig.height=3, fig.width=6}
B = 1000
Tstar = numeric(B)
T1 = mean(col_a)
for (i in 1: B) {
  Xstar = sample(col_a, replace = TRUE)
  Tstar[i] = mean(Xstar) 
}
Tstar15 = quantile(Tstar, 0.015)
Tstar985 = quantile(Tstar, 0.985)
sum(Tstar < Tstar15)
c(2 * T1 - Tstar985, 2 * T1 - Tstar15)
```
With 97% confidence, we can say the mean value is between 5.212 and 6.378

**d)**
We use bootstrap to test the distribution of the sample. we created a list of theta with step 0.01 and test whether it fits the uniform distribution of [3,theta] or not.
```{r, fig.height=3, fig.width=6}
n = length(col_a)
t = max(col_a)
# Function to perform the bootstrap test for a given theta
bootstrap_test <- function(n, t, theta, num_iterations = 1000) {
  tstar = numeric(num_iterations)
  for (i in 1:num_iterations) {
    xstar = runif(n, min = 3, max = theta)
    tstar[i] = max(xstar)
  }
  pl = sum(tstar < t) / num_iterations
  pr = sum(tstar > t) / num_iterations
  p = 2 * min(pl, pr)
}
# Range of theta values to test
theta_values <- seq(3, 12, by = 0.01)  # Adjust the step as needed for precision
# Perform the bootstrap test for each theta
results <- sapply(theta_values, function(theta) bootstrap_test(n, t, theta))
# Find theta values where the proportion of exceeding bootstrap maxima is > alpha (e.g., 0.05)
alpha <- 0.05 # set significance level
theta_not_rejected <- theta_values[results > alpha]
```
```{r echo=FALSE, fig.height=2.5, fig.width=5}
#Plot
plot(theta_values, results, type="l", xlab="Theta", ylab="Proportion Exceeding Observed Max", main = "Bootstrap Test Results")
abline(h=alpha, col="red") #Add horizontal line for alpha level
```

As the figure shows above, we can reject the H0 once theta reaches 7.7.  


We believe Kolmogorov-Smirnov can be applied since we have split the theta into discrete value list, for every theta value KS test is applicable.
```{r, fig.height=3, fig.width=6}
n = length(col_a)
t = max(col_a)
# Range of theta values to test
theta_values <- seq(3, 12, by = 0.1)   
# # Perform the KS test for each theta
results <- sapply(theta_values, function(theta) ks.test(col_a, runif(n, min = 3, max = theta))$p.value)
alpha <- 0.05 # set significance level
theta_not_rejected <- theta_values[results > alpha]
```
```{r echo=FALSE, fig.height=2.5, fig.width=5}
#Plot
plot(theta_values, results, type="l", xlab="Theta", ylab="Proportion Exceeding Observed Max", main = "KS-test Results")
abline(h=alpha, col="red") #Add horizontal line for alpha level
```

This graph showed the KS the p value of KS test. p value is basically rejected when theta is greater than 6.7. After that some low P- value is caused by the stochastic fluctuation since we are measuring the difference of maximum value between our sample and the uniform distribution.


**e)**
To check the median value, we chose the \textbf{binomial test}. if the median is smaller than 6 the number of samples below 6 should be more than 50\%. Here we set median $\geq$ 6 as H0.
```{r, fig.height=3, fig.width=6}
binom.test(sum(col_a < 6), length(col_a), p = 0.5, alternative = "less")
```
it shows that the CI of samples less than 6 is between 0 and 80.1\%. With the P value 0.881, we can not reject the H0. Thus we can not say the median is smaller than 6.
Similarly, we can also use \textbf{binomial test} to test the number of samples smaller than 4.5. However here we set samples < 4.5 more than 25\% as H0.
The output is :
```{r, fig.height=3, fig.width=6}
binom.test(sum(col_a < 4.5), length(col_a), p = 0.25, alternative = "less")
```
Given the output, we can not reject the H0. we can only 95\% confidently say the max portion below 4.5 is 0.377

## Exercise 2 
**a)**
Checking if there is interaction between two factors:
```{r, fig.height=6, fig.width=6}
crops <- read.table("crops.txt", header = TRUE)
crops$County = as.factor(crops$County)
crops$Related = as.factor(crops$Related)
cropsaov_cr1 = lm(Crops ~ County * Related, data = crops)
cropsaov_cr1_anv = anova(cropsaov_cr1)
```

```{r, echo=FALSE, fig.env="figure", fig.pos="!htb"}
tidy_cropsaov_cr1_anv <- broom::tidy(cropsaov_cr1_anv)


knitr::kable(tidy_cropsaov_cr1_anv,
             caption = "ANOVA test of if County and Related influence the crop.",float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")
```

The p value of County:Related is 0.8792 > 0.05, so we do not have enough evidence to reject H0,
which means there is no interaction between County and Related.

Checking whether the 2 factors affect the crops:
```{r, fig.height=6, fig.width=6}
cropsaov_cr2 = lm(Crops ~ Related + County, data = crops)
cropsaov_cr2_drop <-drop1(cropsaov_cr2, test = "F")
```
```{r, echo=FALSE, fig.env="figure", fig.pos="!htb"}
tidy_cropsaov_cr2_drop <- broom::tidy(cropsaov_cr2_drop)


knitr::kable(tidy_cropsaov_cr2_drop,
             caption = "Results of drop1 test of model with Related and County.",float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")
```

The p value of County is 0.4518 > 0.05, so we do not have enough evidence to reject H0, 
which means County does not have a significant effect on the crops.

The p value of Related is 0.5126 > 0.05, so we do not have enough evidence to reject H0, 
which means Related does not have a significant effect on the crops.

Checking model assumptions:
```{r Assignment_2a, echo=FALSE, fig.align='center', fig.cap="Q-Q plot of residuals and residuals fitted plot of model cr1.", fig.pos="H",fig.height=2.5, fig.width=5}

model_data <- augment(cropsaov_cr1)

qq_plot <- ggplot(model_data, aes(sample = .resid)) +
  stat_qq(color = "blue", alpha = 0.7) +
  stat_qq_line(color = "red", linewidth = 1) +
  labs(title = "Q-Q Plot of Residuals",
       x = "Theoretical Quantiles",
       y = "Sample Quantiles") +
  theme_minimal()

resid_fit_plot <- ggplot(model_data, aes(x = .fitted, y = .resid)) +
  geom_point(color = "darkgreen", alpha = 0.7) +
  geom_hline(yintercept = 0, color = "red", linewidth = 1) +
  labs(title = "Residuals vs Fitted Values",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()


combined_plot <- qq_plot + resid_fit_plot
print(combined_plot)

```
The residuals look normal, perhaps a slight curve in the qq-plot.
The plot of the fitted values against the residuals 
shows no pattern, so the assumption of normality is true.

Using the chosen model, estimate the crops 
for a typical farm in County 3 for which landlord and tenant are not related:
```{r, fig.height=6, fig.width=6}
newdata1 <- data.frame(County = factor(3), Related = factor("no"))
print(newdata1)
prediction <- predict(cropsaov_cr2, newdata1)
print(prediction)
```
Since there is no interaction between County and Related, 
we choose the model "cropsaov_cr2" to estimate the result. 
The predicted value is 7760.3 = Intercept (6800.6) + the Estimate of County3 (959.7).
Since both County and Related do not affect crops significantly, the predicted value 
is not very accurate.

**b)**
Whether the influence of Size on Crops is similar for all three counties:

```{r, fig.height=3, fig.width=6}
cropsaov_cs = lm(Crops ~ Related + County * Size, data = crops)
cropsaov_cs_anv <-anova(cropsaov_cs)
cropsaov_cs_smr <-summary(cropsaov_cs)
```
```{r, echo=FALSE, fig.env="figure", fig.pos="!htb"}
tidy_cropsaov_cs_anv <- broom::tidy(cropsaov_cs_anv)
tidy_cropsaov_cs_smr <- broom::tidy(cropsaov_cs_smr)

knitr::kable(tidy_cropsaov_cs_anv,
             caption = "Results of anova test of model with Related + County * Size.",float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")

knitr::kable(tidy_cropsaov_cs_smr,
             caption = "Summary of model with Related + County * Size.",float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")

```

The p-value of County:Size is 0.01192 < 0.05, so we have enough evidence to rejct H0, 
which means there is an interaction between County and Size. 
Furthermore, we take a look at the interaction between each County and Size, 
we can see that County3:Size = 0.17676 > 0.05, which means there is no difference 
between the relation of County1:Size and County3:Size
However, County2:Size = 0.00323 < 0.05, which means there is a difference 
between the relation of County1:Size and County2:Size. 
So the interaction between County and Size does exist.


Checking model assumptions:

```{r, echo=FALSE, fig.align='center', fig.cap="Q-Q plot of residuals and residuals fitted plot of model cs.",  fig.pos="H", fig.height=2.5, fig.width=5, message=FALSE, warning=FALSE}

model_data <- augment(cropsaov_cs)

qq_plot <- ggplot(model_data, aes(sample = .resid)) +
  stat_qq(color = "blue", alpha = 0.7) +
  stat_qq_line(color = "red", linewidth = 1) +
  labs(title = "Q-Q Plot of Residuals",
       x = "Theoretical Quantiles",
       y = "Sample Quantiles") +
  theme_minimal()

resid_fit_plot <- ggplot(model_data, aes(x = .fitted, y = .resid)) +
  geom_point(color = "darkgreen", alpha = 0.7) +
  geom_hline(yintercept = 0, color = "red", linewidth = 1) +
  labs(title = "Residuals vs Fitted Values",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()


combined_plot <- qq_plot + resid_fit_plot
print(combined_plot)

```

The residuals look normal, perhaps a slight curve in the qq-plot.
The plot of the fitted values against the residuals 
shows no pattern, so the assumption of normality can be trust.

whether the influence of Size depends on the relation of landlord and tenant of the farm:
```{r, fig.height=6, fig.width=6}
cropsaov_rs = lm(Crops ~ County + Related * Size, data = crops)
cropsaov_rs_anv <-anova(cropsaov_rs)
```

```{r, echo=FALSE, fig.env="figure", fig.pos="!htb"}
tidy_cropsaov_rs_anv <- broom::tidy(cropsaov_rs_anv)


knitr::kable(tidy_cropsaov_rs_anv,
             caption = "Results of anova test of model with Related and County.",float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")
```

The p-value of Related:Size is 0.29589 > 0.05, we do not have enough evidence to rejct H0,
so there is no interaction between Related and Size.

Checking model assumptions:
```{r, echo=FALSE, fig.align='center', fig.cap="Q-Q plot of residuals and residuals fitted plot of model rs.",  fig.pos="H",fig.height=2.5, fig.width=5, message=FALSE, warning=FALSE}

model_data <- augment(cropsaov_rs)

qq_plot <- ggplot(model_data, aes(sample = .resid)) +
  stat_qq(color = "blue", alpha = 0.7) +
  stat_qq_line(color = "red", linewidth = 1) +
  labs(title = "Q-Q Plot of Residuals",
       x = "Theoretical Quantiles",
       y = "Sample Quantiles") +
  theme_minimal()

resid_fit_plot <- ggplot(model_data, aes(x = .fitted, y = .resid)) +
  geom_point(color = "darkgreen", alpha = 0.7) +
  geom_hline(yintercept = 0, color = "red", linewidth = 1) +
  labs(title = "Residuals vs Fitted Values",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()


combined_plot <- qq_plot + resid_fit_plot
print(combined_plot)

```

The QQ-normal plot shows normality is doubtful. Also the spread in the residuals seems 
to be bigger for smaller fitted values. Some data-points also seem extreme. 
So the model also cannot be trusted.

Overall, based on our analysis above, we should choose the model 'cropsaov_cs'with Related + County * Size.

**c)**
For the resulting model from b), investigate how County, Related and Size influence Crops.

```{r, fig.height=6, fig.width=6}
cropsaov_cs = lm(Crops ~ Related + County * Size, data = crops)

#test: Related
cropsaov_cs2 = lm(Crops ~ County * Size, data = crops)
cropsaov_cs2_anv <-anova(cropsaov_cs2, cropsaov_cs)
```

```{r, echo=FALSE, fig.env="figure", fig.pos="!htb"}
tidy_cropsaov_cs2_anv <- broom::tidy(cropsaov_cs2_anv)


knitr::kable(tidy_cropsaov_cs2_anv,
             caption = "Results of anova test of model with Size and County.",float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")
```

The p-value is 0.4988 > 0.05, which means there is no significant difference between 2 models,
so factor Related does not affect crops.

Checking model assumptions:
```{r, echo=FALSE, fig.align='center', fig.cap="Q-Q plot of residuals and residuals fitted plot of model cs2.",, fig.pos="H", fig.height=2, fig.width=4}

model_data <- augment(cropsaov_cs2)

qq_plot <- ggplot(model_data, aes(sample = .resid)) +
  stat_qq(color = "blue", alpha = 0.7) +
  stat_qq_line(color = "red", linewidth = 1) +
  labs(title = "Q-Q Plot of Residuals",
       x = "Theoretical Quantiles",
       y = "Sample Quantiles") +
  theme_minimal()

resid_fit_plot <- ggplot(model_data, aes(x = .fitted, y = .resid)) +
  geom_point(color = "darkgreen", alpha = 0.7) +
  geom_hline(yintercept = 0, color = "red", linewidth = 1) +
  labs(title = "Residuals vs Fitted Values",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()


combined_plot <- qq_plot + resid_fit_plot
print(combined_plot)
```
The residuals look normal, perhaps a slight curve in the qq-plot.
The plot of the fitted values against the residuals 
shows no pattern, so the assumption of normality is true.

```{r, fig.height=6, fig.width=6}
#test: County
cropsaov_rs2 = lm(Crops ~ Related + Size + County:Size, data = crops)
cropsaov_rs2_anv <- anova(cropsaov_rs2, cropsaov_cs)
```
```{r, echo=FALSE, fig.env="figure", fig.pos="!htb"}
tidy_cropsaov_rs2 <- broom::tidy(cropsaov_rs2)


knitr::kable(tidy_cropsaov_rs2,
             caption = "Results of anova test of model with Size and County and the interaction country and size.",float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")
```
```{r, echo=FALSE, fig.align='center', fig.cap="Q-Q plot of residuals and residuals fitted plot of model rs2.", fig.pos="H",fig.height=2, fig.width=4, message=FALSE, warning=FALSE}

model_data <- augment(cropsaov_rs2)

qq_plot <- ggplot(model_data, aes(sample = .resid)) +
  stat_qq(color = "blue", alpha = 0.7) +
  stat_qq_line(color = "red", linewidth = 1) +
  labs(title = "Q-Q Plot of Residuals",
       x = "Theoretical Quantiles",
       y = "Sample Quantiles") +
  theme_minimal()

resid_fit_plot <- ggplot(model_data, aes(x = .fitted, y = .resid)) +
  geom_point(color = "darkgreen", alpha = 0.7) +
  geom_hline(yintercept = 0, color = "red", linewidth = 1) +
  labs(title = "Residuals vs Fitted Values",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()


combined_plot <- qq_plot + resid_fit_plot
print(combined_plot)
```


The p-value is 0.02551 < 0.05, which means there is significant difference between 2 models,
so factor County itself does affect crops significantly. Also from b), we know that there is a relation 
between County and Size.

Checking model assumptions:
```{r, echo=FALSE, fig.align='center', fig.cap="Q-Q plot of residuals and residuals fitted plot of model rs2.", fig.pos="H", fig.height=2, fig.width=4, message=FALSE, warning=FALSE}

model_data <- augment(cropsaov_rs2)

qq_plot <- ggplot(model_data, aes(sample = .resid)) +
  stat_qq(color = "blue", alpha = 0.7) +
  stat_qq_line(color = "red", linewidth = 1) +
  labs(title = "Q-Q Plot of Residuals",
       x = "Theoretical Quantiles",
       y = "Sample Quantiles") +
  theme_minimal()

resid_fit_plot <- ggplot(model_data, aes(x = .fitted, y = .resid)) +
  geom_point(color = "darkgreen", alpha = 0.7) +
  geom_hline(yintercept = 0, color = "red", linewidth = 1) +
  labs(title = "Residuals vs Fitted Values",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()


combined_plot <- qq_plot + resid_fit_plot
print(combined_plot)
```
The residuals look normal, perhaps a slight curve in the qq-plot.
The plot of the fitted values against the residuals 
shows no pattern, so the assumption of normality is true.

```{r, fig.height=6, fig.width=6}
#test: Size
cropsaov_crs = lm(Crops ~ Related + County + County:Size, data = crops)
cropsaov_crs_anv <- anova(cropsaov_crs, cropsaov_cs)

```
```{r, echo=FALSE, fig.env="figure", fig.pos="!htb"}
tidy_cropsaov_crs_anv <- broom::tidy(cropsaov_crs_anv)


knitr::kable(tidy_cropsaov_crs_anv,
             caption = "Results of anova test of model with Related and County and the interaction country and size.",float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")
```


The p-value is 0 (which is significant small) < 0.05, which means there is significant difference between 2 models,
so factor Size itself does affect crops significantly. Also from b), we know that there is a relation 
between County and Size.

Checking model assumptions:
```{r, echo=FALSE, fig.align='center', fig.cap="Q-Q plot of residuals and residuals fitted plot of model crs.", fig.pos="H", fig.height=2.5, fig.width=5, message=FALSE, warning=FALSE}

model_data <- augment(cropsaov_crs)

qq_plot <- ggplot(model_data, aes(sample = .resid)) +
  stat_qq(color = "blue", alpha = 0.7) +
  stat_qq_line(color = "red", linewidth = 1) +
  labs(title = "Q-Q Plot of Residuals",
       x = "Theoretical Quantiles",
       y = "Sample Quantiles") +
  theme_minimal()

resid_fit_plot <- ggplot(model_data, aes(x = .fitted, y = .resid)) +
  geom_point(color = "darkgreen", alpha = 0.7) +
  geom_hline(yintercept = 0, color = "red", linewidth = 1) +
  labs(title = "Residuals vs Fitted Values",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()


combined_plot <- qq_plot + resid_fit_plot
print(combined_plot)
```
The residuals look normal, perhaps a slight curve in the qq-plot.
The plot of the fitted values against the residuals 
shows no pattern, so the assumption of normality is true.

**d)**
Using the resulting model from b), predict the crops for a farm from County 2 of size 165, 
with related landlord and tenant. Estimate also the error variance.
```{r, fig.height=6, fig.width=6}
newdata2 <- data.frame(County = factor(2), Related = factor("yes"), Size = 165)
print(newdata2)
prediction <- predict(cropsaov_cs, newdata2)
prediction <- predict(cropsaov_cs, newdata2)
print(prediction)

```
So, the crops for a farm from County 2 of size 165, with related landlord and tenant should be 6141.
Table 7 shows the estimation of  error variance.

## Exercise 3
**a)**
```{r, fig.height=3, fig.width=6}
# Define the blocks and treatments
blocks <- 1:6
plots_per_block = 4
if_treated <- c(1, 1, 0, 0)
# Create a function to randomize treatments for a single block
randomize_block <- function() {
  sample(if_treated)
}
# Apply randomization to each block
randomized_design <- data.frame(
  blocks = rep(blocks, each = 4),  # 4 plots per block
  plot = 1:plots_per_block,
  N = unlist(lapply(blocks, function(x) randomize_block())),
  P = unlist(lapply(blocks, function(x) randomize_block())),
  K = unlist(lapply(blocks, function(x) randomize_block()))
)
```

**b)**

```{r include=FALSE}
library(MASS)
data(npk)
library(dplyr)
library(ggplot2)

```

```{r, fig.height=2.5, fig.width=5}

avg_yield <- npk %>%
  group_by(block, N) %>%
  summarise(mean_yield = mean(yield), .groups = 'drop')

ggplot(avg_yield, aes(x = block, y = mean_yield, color = N, group = N)) +
  geom_line() +
  geom_point() +
  labs(title = "Average Yield by Block and Nitrogen Treatment",
       x = "Block", y = "Average Yield")
```


The plot shows that yields with nitrogen (N=1) are generally higher across all blocks, indicating a positive effect of nitrogen. However, variability exists between blocks,with the role of N not being as pronounced in block 5, whereas it was very significant in blocks 1 and 4


**c)**
```{r, fig.height=3, fig.width=6}
block = as.factor(npk$block)
N = as.factor(npk$N)
P = as.factor(npk$P)
K = as.factor(npk$K)
npkaov_bn = lm(yield ~ block * N, data = npk)
npkaov_bn_anv <-anova(npkaov_bn)
```
```{r, echo=FALSE, fig.env="figure", fig.pos="!htb"}
tidy_npkaov_bn_anv <- broom::tidy(npkaov_bn_anv)


knitr::kable(tidy_npkaov_bn_anv,
             caption = "Results of anova test of model block*N.",float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")
```
As the result shown, the interaction block:N is insignificant, suggesting no block-specific nitrogen effects.
Then The model can be simplified as:
```{r, fig.height=3, fig.width=6}
npkaov_bn = lm(yield ~ block + N, data = npk)
npkaov_bn_anv <- anova(npkaov_bn)
```
```{r, echo=FALSE, fig.env="figure", fig.pos="!htb"}
tidy_npkaov_bn_anv <- broom::tidy(npkaov_bn_anv)


knitr::kable(tidy_npkaov_bn_anv,
             caption = "Results of anova test of model block+N.",float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")
```
Base in the result of anova, Block effect is significant, justifying its inclusion to control for variability between blocks.
Meanwhile,N effect is significant, indicating nitrogen boosts yield.

The let's check the model assumptions:

```{r, echo=FALSE, fig.align='center', fig.cap="Q-Q plot of residuals and residuals fitted plot of model crs.",  fig.pos="H", fig.height=2.5, fig.width=5, message=FALSE, warning=FALSE}

model_data <- augment(npkaov_bn)

qq_plot <- ggplot(model_data, aes(sample = .resid)) +
  stat_qq(color = "blue", alpha = 0.7)  +
  labs(title = "Q-Q Plot of Residuals",
       x = "Theoretical Quantiles",
       y = "Sample Quantiles") +
  theme_minimal()

resid_fit_plot <- ggplot(model_data, aes(x = .fitted, y = .resid)) +
  geom_point(color = "darkgreen", alpha = 0.7) +
  geom_hline(yintercept = 0, color = "red", linewidth = 1) +
  labs(title = "Residuals vs Fitted Values",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()


combined_plot <- qq_plot + resid_fit_plot
print(combined_plot)
```

The QQ-normal plot is slight curved but overall normal. Also the residuals are evenly distributed above and below the red line

So, including block is sensible because it explains significant variability, improving model accuracy. Ignoring blocks could lead to biased estimates of the nitrogen effect.


The Friedman test cannot be used in this situation. The data here is not a repeated measures design: 
The Friedman test can only be used when the same subjects are measured under different conditions.  
Each block here has different plots, so we are not measuring the same plot across different fertilizer combinations.

**d)**

```{r, fig.height=3, fig.width=6}
# Basic model without interactions, main effects only
model0 <- lm(yield ~ block + N + P + K , data = npk)
# Models with one block interaction term
model1 <- lm(yield ~ block + N + P + K + block:N, data = npk)
model2 <- lm(yield ~ block + N + P + K + block:P, data = npk)
model3 <- lm(yield ~ block + N + P + K + block:K, data = npk)

# Check significance of interactions
anv1 <- anova(model0,model1)
anv2 <- anova(model0,model2) 
anv3 <- anova(model0,model3) 
```

```{r, echo=FALSE, fig.env="figure", fig.pos="!htb"}
tidy_anv1 <- broom::tidy(anv1)
tidy_anv2 <- broom::tidy(anv2)
tidy_anv3 <- broom::tidy(anv3)

knitr::kable(tidy_anv1,
             caption = "Anova test of base model and models with one block interaction term.",
             float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")
knitr::kable(tidy_anv2,
             float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")
knitr::kable(tidy_anv3,
             float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")
```
The results show that all interactions are not significant.Then we move on to examine the significance of each main effect:

```{r, fig.height=3, fig.width=6}
model0_drop <- drop1(model0,test="F")
```
```{r, echo=FALSE, fig.env="figure", fig.pos="!htb"}
tidy_model0_drop <- broom::tidy(model0_drop)


knitr::kable(tidy_model0_drop,
             caption = "Results of drop test of model block+N+K+P.",float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")
```

The anova and AIC analysis showed that the main effects block and N,K contributed significantly to the model, while P was  insignificant.
Thus, our final model is *yield=block + N + K* :
```{r, fig.height=3, fig.width=6}
model_final <- lm(yield ~ block + N + K , data = npk)
model_final_anv <-anova(model_final)
```
```{r, echo=FALSE, fig.env="figure", fig.pos="!htb"}
tidy_model_final_anv <- broom::tidy(model_final_anv)


knitr::kable(tidy_model_final_anv,
             caption = "Results of anova test of model block+N+K.",float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")
```
All main effect contributions in this model are significant and increase in significance over all of the above models.




**e)**
```{r include=FALSE}
library(emmeans)
```

```{r, fig.height=3, fig.width=6}
model_final_N_emm <- emmeans(model_final, specs = ~ N)  
model_final_K_emm <- emmeans(model_final, specs = ~ K)  
model_final_b_emm <- emmeans(model_final, specs = ~ block)  
```

```{r, echo=FALSE, fig.env="figure", fig.pos="!htb"}
tidy_model_final_N_emm <- broom::tidy(model_final_N_emm)
tidy_model_final_K_emm <- broom::tidy(model_final_K_emm)
tidy_model_final_b_emm <- broom::tidy(model_final_b_emm)

knitr::kable(tidy_model_final_N_emm,
             caption = "Compute estimated marginal means (EMMs) for N,K,block separately.",
             float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")
knitr::kable(tidy_model_final_K_emm,
             float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")
knitr::kable(tidy_model_final_b_emm,
             float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")
```

According to the results, it can be concluded that applying N will make the yield rise and applying P will make the yield fall, and different blocks have different base yields, from smallest to largest are: block4,block5,block1,block6,block2,block3.
The combination leading to the maximum yield should be: in block3, apply nitrogen (N=1) and not apply potassium (K=0).P is not included in the final model so it is not discussed.

**f)**
```{r include=FALSE}
library(lme4)
```

```{r, fig.height=3, fig.width=6}
mixed_model <- lmer(yield ~ N +K +(1|block), data = npk,REML=FALSE)
mixed_model_K <- lmer(yield ~ K +(1|block), data = npk,REML=FALSE)
mixed_anv <- anova(mixed_model_K,mixed_model)
```
```{r, echo=FALSE, fig.env="figure", fig.pos="!htb"}
tidy_mixed_anv <- broom::tidy(mixed_anv)


knitr::kable(tidy_mixed_anv,
             caption = "ANOVA for mixed eﬀects model with and without N.",float=FALSE,
             format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "5cm")
```
Through test of the mixed-effects model, the addition of N significantly improved the model's ability to explain yield (p = 0.001), and the model complexity was reasonably controlled (AIC/BIC decreased).Thus we can conclude that nitrogen ( N ) has a significant effect on yield.
In contrast to the results of fixed eﬀects model (Table 17), the p-value of N in the anova analysis of the fixed effects model is 0.003, again significant but not as significant as the mixed effects model. 
So we can get the conclusion:Nitrogen significantly increases yield, validated by both fixed and mixed models.